### 2023-12-17 tesis_Ashish/Chapter4/chapter4.tex
Sehrawat A thesis 17 Dec 2023

For studying the performance of PCC luminometer during 2018 data taking, we have divided CMS luminosity data taking into seven periods: A, B, C, D1, D2, D3, and D4, as depicted in % 2018 CMS luminosity data taking is divided into seven periods namely A, B, C, D1, D2, D3 and D4
Fig. \ref{fig:period_bound}. Run ranges for each period are shown in Table \ref{tab:period_run_ranges}.
During period B, specifically corresponding to Fill 6868, van der Meer (vdM) scans were conducted for calibrating the luminometer.
%vdM data is shown during period B corresponding to Fill 6868.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/period_boundary.png}
\caption[2018 CMS luminosity data taking periods.]{%
   2018 luminosity data taking periods showing data taking period boundaries and vdM calibration data  \cite{CERNLumiPublicResults}.
}
\label{fig:period_bound}
\end{figure}

\begin{table}
  \begin{center}
    \begin{tabular}{ccccc}
    \textbf{Period}   & \textbf{Run range} \\ \hline
     2018A      &        315252-316995         \\
     2018B      &        317080-319311         \\
     2018C      &        319337-320065         \\
     2018D1     &        320500-321665        \\
     2018D2     &        321710-322964         \\
     2018D3     &        323363-324420         \\
     2018D4     &        324564-325175        \\
      \end{tabular}
    \caption[Run ranges]{2018 luminosity data periods with run ranges.}
    \label{tab:period_run_ranges}
  \end{center}
\end{table}

In the study under consideration, we have utilized a series of datasets corresponding to the data-taking periods of 2018, namely 2018A, 2018B, 2018C, and 2018D as shown in Table \ref{tab:datasets}. These datasets have been drawn from the Alignment and Calibration (AlCa) project, an integral component of the Compact Muon Solenoid (CMS) experiment conducted at the Large Hadron Collider. The dataset selection includes two subsets denoted as AlCaPCCRandom and AlCaPCCZeroBias. The AlCaPCCRandom dataset comprises of data where trigger is applied on both colliding and non-colliding bunches, thereby providing a holistic and representative view of the whole dataset. On the other hand, the AlCaPCCZeroBias dataset encapsulates events that have passed the trigger system with no bias where trigger is applied on only colliding bunches, thereby serving as an unbiased reference dataset. In terms of processing, the selected datasets have undergone an Express processing pathway. This rapid processing approach, while less detailed than the full reprocessing, provides valuable and timely insights into the data quality and physics performance. The datasets fall within the ALCARECO (ALignment, CALibration, and REConstruction objects) data tier, which specifically encompasses information pertinent to detector alignment and calibration, forming a crucial resource in our analysis. The timestamps associated with these datasets (e.g., $'13Mar2023\_UL2018\_PCC'$) indicate that they have undergone a reprocessing or update post their initial collection in 2018.
%This reprocessing phase enhances the scientific value of these datasets, by incorporating improved calibrations and alignment, and updates to the data reconstruction algorithms. Consequently, the reprocessed datasets used in this thesis afford a more accurate and comprehensive representation of the physics events under investigation.

@@

%\textbf{/AlCaLumiPixels/Run2018A-AlCaPCCRandom15Apr2023\_UL2018\_315252\_316062\_PCCRandom\_-v1/ALCARECO}

\textbf{\url{/AlCaLumiPixels/Run2018A-AlCaPCCRandom15Apr2023_UL2018_315252_316062_PCCRandom_-v1/ALCARECO}}
%\textbf{/AlCaLumiPixels/Run2018A-AlCaPCCRandom15Apr2023_UL2018_315252_316062_PCCRandom_-v1/ALCARECO}

@@

\item /AlCaLumiPixels/Run2018A: This indicates that the dataset is related to the "AlCa" (ALignment and CALibration) stream, specifically for the "LumiPixels" subdetector. The data corresponds to the Run2018A period, which means it was collected during the first run period in 2018.

\item AlCaPCCRandom: This tag represents the primary dataset and the trigger path used to collect the data. "PCC" stands for "Pixel Cluster Counting" and is a method for luminosity measurement. "Random" indicates that this dataset was collected using a random trigger, which is a method to sample a fraction of the events by applying trigger on both colliding and non-colliding bunch crossings.

\item $15Apr2023\_UL2018\_315252\_316062\_PCCRandom\_-v1$: This part represents the processing version, date, and run range of the dataset. "15Apr2023" is the date when this specific version of the dataset was created. "UL2018" indicates that this data has been processed using the "Ultra Legacy" (UL) campaign for the 2018 dataset, which aims to provide the highest quality data for physics analysis. The numbers $"315252_316062"$ represent the range of run numbers included in the dataset. $"PCCRandom_"$ is a reference to the Pixel Cluster Counting method used for luminosity measurement and the random trigger. "v1" denotes the version number of this processed dataset.

\item ALCARECO:
This acronym stands for "ALignment and CALibration RECOstruction." It is a data tier representing the output of specific reconstruction algorithms applied to the raw data to produce high-level objects needed for the alignment and calibration tasks.

@@

%To summarize, this dataset is a collection of events from the CMS experiment during the Run2018A period, sampled using a random trigger. It is specifically related to the luminosity measurement using the Pixel Cluster Counting method. The dataset has been processed using the Ultra Legacy campaign, and the output contains high-level objects required for alignment and calibration tasks.

%$\textbf{/StreamALCALUMIPIXELSEXPRESS/Run2018A-AlCaPCCRandom-Express-v1/ALCARECO}$

\textbf{\url{/StreamALCALUMIPIXELSEXPRESS/Run2018A-AlCaPCCRandom-Express-v1/ALCARECO}}

\item /StreamALCALUMIPIXELSEXPRESS/Run2018A:
This indicates that the dataset is related to the "AlCa" (ALignment and CALibration) stream, specifically for the "LumiPixels" subdetector, and is part of the "Express" data processing tier. The data corresponds to the Run2018A period, which means it was collected during the first run period in 2018.

\item AlCaPCCRandom:
This tag represents the primary dataset and the trigger path used to collect the data. "PCC" stands for "Pixel Cluster Counting" and is a method for luminosity measurement. "Random" indicates that this dataset was collected using a random trigger, which is a method to sample a fraction of the events without specific selection criteria (i.e., unbiased events).

\item Express-v1:
This part represents the processing version and the type of data processing. "Express" indicates that this data has been processed using the "Express" campaign, which is a near-real-time processing of the data as it is collected. This type of processing is essential for providing timely feedback to the experiments and enabling rapid identification of any issues. "v1" denotes the version number of this processed dataset.

\item ALCARECO:
This acronym stands for "ALignment and CALibration RECOstruction." It is a data tier representing the output of specific reconstruction algorithms applied to the raw data to produce high-level objects needed for the alignment and calibration tasks.

%In summary, this dataset is a collection of events from the CMS experiment during the Run2018A period, sampled using a random trigger. It is specifically related to the luminosity measurement using the Pixel Cluster Counting method. The dataset has been processed using the Express campaign, and the output contains high-level objects required for alignment and calibration tasks.

%$\textbf{/AlCaLumiPixels/Run2018A-AlCaPCCZeroBias-13Mar2023_UL2018_PCC-v1/ALCARECO}$

%$\textbf{/AlCaLumiPixels/Run2018A-AlCaPCCZeroBias-13Mar2023\_UL2018\_PCC-v1/ALCARECO}$

\textbf{\url{/AlCaLumiPixels/Run2018A-AlCaPCCZeroBias-13Mar2023_UL2018_PCC-v1/ALCARECO}}

\item /AlCaLumiPixels/Run2018A:
This indicates that the dataset is related to the "AlCa" (ALignment and CALibration) stream, specifically for the "LumiPixels" subdetector. The data corresponds to the Run2018A period, which means it was collected during the first run period in 2018.

\item AlCaPCCZeroBias:
This tag represents the primary dataset and the trigger path used to collect the data. "PCC" stands for "Pixel Cluster Counting" and is a method for luminosity measurement. "ZeroBias" indicates that this dataset was collected using a trigger with no specific selection criteria (i.e., unbiased events).

\item $13Mar2023_UL2018_PCC-v1$:
This part represents the processing version and the date of the dataset. "13Mar2023" is the date when this specific version of the dataset was created. "UL2018" indicates that this data has been processed using the "Ultra Legacy" (UL) campaign for the 2018 dataset, which aims to provide the highest quality data for physics analysis. "PCC" is again a reference to the Pixel Cluster Counting method used for luminosity measurement. "v1" denotes the version number of this processed dataset.

\item ALCARECO:
This acronym stands for "ALignment and CALibration RECOstruction." It is a data tier representing the output of specific reconstruction algorithms applied to the raw data to produce high-level objects needed for the alignment and calibration tasks.

%So, to summarize, this dataset is a collection of unbiased events from the CMS experiment during the Run2018A period. It is specifically related to the luminosity measurement using the Pixel Cluster Counting method. The dataset has been processed using the Ultra Legacy campaign, and the output contains high-level objects required for alignment and calibration tasks.

%$\textbf{/AlCaLumiPixels/Run2018D-AlCaPCCZeroBias-PromptReco-v2/ALCARECO}$

\textbf{\url{/AlCaLumiPixels/Run2018D-AlCaPCCZeroBias-PromptReco-v2/ALCARECO}}

\item /AlCaLumiPixels/Run2018D:
This indicates that the dataset is related to the "AlCa" (ALignment and CALibration) stream, specifically for the "LumiPixels" subdetector. The data corresponds to the Run2018D period, which means it was collected during the fourth run period in 2018.

\item AlCaPCCZeroBias:
This tag represents the primary dataset and the trigger path used to collect the data. "PCC" stands for "Pixel Cluster Counting" and is a method for luminosity measurement. "ZeroBias" indicates that this dataset was collected using a trigger with no specific selection criteria (i.e., unbiased events).

\item PromptReco-v2:
This part represents the processing version and the type of data processing. "PromptReco" indicates that this data has been processed using the "Prompt Reconstruction" campaign, which is a near-real-time processing of the data as it is collected. This type of processing is essential for providing timely feedback to the experiments and enabling rapid identification of any issues. "v2" denotes the version number of this processed dataset.

\item ALCARECO:
This acronym stands for "ALignment and CALibration RECOstruction." It is a data tier representing the output of specific reconstruction algorithms applied to the raw data to produce high-level objects needed for the alignment and calibration tasks.

%In summary, this dataset is a collection of unbiased events from the CMS experiment during the Run2018D period, specifically related to the luminosity measurement using the Pixel Cluster Counting method. The dataset has been processed using the Prompt Reconstruction campaign, and the output contains high-level objects required for alignment and calibration tasks.

The dataset /StreamALCALUMIPIXELSEXPRESS/Run2018A-AlCaPCCRandom-Express-v1/ALCARECO, is a dataset from the CMS experiment at CERN's Large Hadron Collider (LHC). This dataset is specifically collected for luminosity measurements using pixel cluster counting (PCC) in the CMS detector by applying trigger on colliding and non-colliding bunches.

\item StreamALCALUMIPIXELSEXPRESS: This indicates that the dataset is related to the "AlCa" (Alignment and Calibration) sub-system of the CMS detector, specifically focused on luminosity measurements using the pixel detector. The "EXPRESS" in the name signifies that this dataset is processed using the Express Stream, which is a fast data processing system for near real-time analysis of CMS data.

\item Run2018A: This represents the run period during which the data was collected. In this case, it refers to the 2018 data-taking period, specifically the "A" run.

\item AlCaPCCRandom: This indicates that the dataset is related to the "PCC" (Pixel Cluster Counting) method of luminosity measurement, and "Random" refers to the trigger used for data collection. In this case, a random trigger is used, selecting events without any specific bias.

\item Express-v1: This signifies that the dataset has been processed using the Express Stream system. The "-v1" indicates the version of the processing.

\item ALCARECO: This is the final data format of the dataset, specifically designed for alignment and calibration tasks.

%\item Data collection: The CMS detector collects data from proton-proton collisions at the LHC. For this dataset, a random trigger is used , ensuring a representative sample of events for luminosity measurements.

%\item Data processing: The raw data is processed using the Express Stream system. This is a fast processing system used for near real-time analysis of CMS data. During this process, the raw data is converted into higher-level objects, such as tracks and vertices, which are more convenient for analysis.

%\item Data filtering: ALCARECO datasets are created by applying specific filters and selections to the processed data. In this case, the filters are chosen to focus on pixel cluster counting for luminosity measurements.

%\item Dataset storage: The final ALCARECO dataset is stored and made available for analysis by researchers working on alignment, calibration, and luminosity-related studies.

 The dataset /AlCaLumiPixels/Run2018D-AlCaPCCZeroBias-PromptReco-v2/ALCARECO, is specifically collected for luminosity measurements using pixel cluster counting (PCC) in the CMS detector by applying trigger on only colliding bunches.

\item AlCaLumiPixels: This indicates that the dataset is related to the "AlCa" (Alignment and Calibration) sub-system of the CMS detector, specifically focused on luminosity measurements using the pixel detector.

\item Run2018A: This represents the run period during which the data was collected. In this case, it refers to the 2018 data-taking period, specifically the "A" run.

\item AlCaPCCZeroBias: This indicates that the dataset is related to the "PCC" (Pixel Cluster Counting) method of luminosity measurement, and "ZeroBias" refers to the trigger used for data collection. ZeroBias triggers are unbiased, meaning they are randomly triggered and do not rely on any specific event characteristics.

\item PromptReco-v2: This signifies that the dataset has been processed using the "PromptReco" (Prompt Reconstruction) system, which is a fast processing system for real-time analysis of CMS data. The "-v2" indicates the version of the processing.

\item ALCARECO: This is the final data format of the dataset, specifically designed for alignment and calibration tasks.

%\item Data collection: The CMS detector collects data from proton-proton collisions at the LHC. For this dataset, the ZeroBias trigger is used, ensuring an unbiased sample of events for luminosity measurements.

%\item Data processing: The raw data is processed using the Prompt Reconstruction system. During this process, the raw data is converted into higher-level objects, such as tracks and vertices, which are more convenient for analysis.

%\item Data filtering: ALCARECO datasets are created by applying specific filters and selections to the processed data. In this case, the filters are chosen to focus on pixel cluster counting for luminosity measurements.

%\item Dataset storage: The final ALCARECO dataset is stored and made available for analysis by researchers working on alignment, calibration, and luminosity-related studies.

The dataset "AlCaLumiPixels/Run2018D-AlCaPCCZeroBias-PromptReco-v2/ALCARECO" dataset is a specialized dataset that provides information related to luminosity measurements using pixel detectors.

\item AlCaLumiPixels: The term "AlCaLumiPixels" indicates that the dataset focuses on luminosity measurements obtained from the pixel detectors. The pixel detectors are part of the CMS experiment's innermost layers and are designed to precisely track the paths of charged particles produced in proton-proton collisions.

\item Run2018D: The "Run2018D" refers to a specific data-taking period or run period of the CMS experiment. The LHC operates in cycles of data-taking known as runs, which are typically identified by a year and a letter. In this case, "2018D" corresponds to a particular period of data collection in 2018.

\item AlCaPCCZeroBias: "AlCaPCCZeroBias" is a term that likely indicates the specific algorithm and event type used for luminosity measurement. The "PCC" stands for "Pixel Cluster Counting," which is an algorithm employed to estimate the instantaneous luminosity based on the information recorded by the pixel detectors. The "ZeroBias" events represent a special category of collisions where all the CMS detector systems trigger, regardless of the specific physics processes involved. These events are particularly useful for luminosity measurements.

\item PromptReco-v2: "PromptReco" refers to the reconstruction process for the prompt data. Prompt reconstruction involves the real-time processing and calibration of the recorded data immediately after it is taken. The "-v2" indicates a specific version or iteration of the prompt reconstruction process.

\item ALCARECO: "ALCARECO" stands for "ALignment CALibration RECOnditioned" data. It is a category of specialized reconstructed datasets produced with specific requirements for alignment and calibration studies. These datasets are carefully processed and conditioned to ensure accurate alignment of the CMS detector elements and precise calibration of the recorded data.

PromptReco (Prompt Reconstruction) and AlCaReco (ALignment and CALibration REConstructed Objects) are two different data processing and formatting systems used in the CMS experiment at CERN's Large Hadron Collider (LHC). While both systems are related to the processing of data collected by the CMS detector, they serve different purposes and have distinct features. Here is a comparison of the two systems:

Purpose: \\ 
PromptReco: The primary goal of the PromptReco system is to rapidly convert the raw data collected by the CMS detector into a format more suitable for physics analysis. It reconstructs the raw data into higher-level objects like tracks, vertices, and calorimeter clusters, which are then used by researchers to study various aspects of the particle collisions.

AlCaReco: The main purpose of the AlCaReco system is to provide specialized datasets for detector alignment, calibration, and other related tasks. AlCaReco datasets are generated by applying specific filters and selections to the reconstructed data (RECO), focusing on events and information that are particularly useful for alignment, calibration, and performance monitoring.

Data format: \\
PromptReco: The output of the PromptReco system is the RECO (REConstructed Objects) data format. This is the primary data format used for physics analysis in the CMS experiment, containing reconstructed objects such as tracks, vertices, and calorimeter clusters.

AlCaReco: The AlCaReco data format is designed specifically for alignment, calibration, and performance monitoring tasks. It is a more focused and compact data format, containing only the events and information relevant to the specific task, resulting in reduced data volume compared to the full RECO data.

Data processing:\\
PromptReco: The PromptReco system processes the raw data at Tier-0 of the Worldwide LHC Computing Grid (WLCG), using a series of algorithms to reconstruct the detector information into higher-level physics objects.

AlCaReco: The AlCaReco processing stream is applied after the PromptReco system, using the reconstructed data (RECO) as input. AlCaReco datasets are generated by applying specific filters and selections tailored to the needs of the alignment, calibration, and performance monitoring tasks.

%The main difference between PromptReco and AlCaReco is their purpose and the type of data they produce. PromptReco is focused on the rapid reconstruction of raw data into a format suitable for physics analysis, while AlCaReco is designed to provide specialized datasets for detector alignment, calibration, and performance monitoring tasks. Both systems play a crucial role in the overall data processing and analysis pipeline of the CMS experiment.

%AlCaReco stands for "ALignment and CALibration REConstructed Objects." It is a specific data format and processing stream used by the CMS experiment at CERN's Large Hadron Collider (LHC) for the purpose of detector alignment, calibration, and other specialized tasks. AlCaReco datasets are designed to be smaller and more focused compared to the primary reconstructed data (RECO), making them suitable for specialized analyses that require a specific set of event characteristics or detector information. The AlCaReco data format and processing stream serve several important functions: Customized event selection: AlCaReco datasets are generated by applying specific filters and selections to the reconstructed data (RECO). These filters are tailored to select events that are particularly useful for alignment, calibration, and other specialized tasks. For example, certain triggers or event topologies might be more suitable for studying the performance of a specific detector subsystem or measuring certain quantities. By selecting events with these specific characteristics, the AlCaReco datasets provide a more focused sample for the targeted analyses. Reduced data volume: One of the main advantages of AlCaReco datasets is their reduced data volume compared to the full RECO data. By selecting only the events and information relevant to the specific task, the AlCaReco data format helps save storage space and computational resources, making the data processing and analysis more efficient. Detector performance monitoring: AlCaReco datasets are used to monitor the performance of the CMS detector and its subsystems. By analyzing the data in these specialized datasets, researchers can track the detector's performance over time and identify any potential issues or areas for improvement. This information can then be used to optimize the detector's operation and ensure accurate data collection. Alignment and calibration: Accurate detector alignment and calibration are crucial for the overall performance of the CMS experiment. AlCaReco datasets provide the necessary information to perform these tasks, allowing researchers to refine the detector's geometry, timing, and energy response. This, in turn, improves the quality of the data collected by the CMS detector and increases the accuracy of the physics analyses. In summary, AlCaReco is a specialized data format and processing stream used by the CMS experiment for tasks related to detector alignment, calibration, and performance monitoring. By providing focused datasets with reduced data volume, AlCaReco enables efficient and accurate analysis of the data needed to maintain and optimize the performance of the CMS detector. AlCaReco: The main purpose of the AlCaReco system is to provide specialized datasets for detector alignment, calibration, and other related tasks. AlCaReco datasets are generated by applying specific filters and selections to the reconstructed data (RECO), focusing on events and information that are particularly useful for alignment, calibration, and performance monitoring. AlCaReco: The AlCaReco data format is designed specifically for alignment, calibration, and performance monitoring tasks. It is a more focused and compact data format, containing only the events and information relevant to the specific task, resulting in reduced data volume compared to the full RECO data. AlCaReco: The AlCaReco processing stream is applied after the PromptReco system, using the reconstructed data (RECO) as input. AlCaReco datasets are generated by applying specific filters and selections tailored to the needs of the alignment, calibration, and performance monitoring tasks.

The calibration of luminosity measurement by the CMS experiment luminometers in the 2018 proton-proton data taking at $\sqrt{s}$ = 13 TeV \cite{CMS-PAS-LUM-18-002} was performed during

\begin{itemize}
\item LHC fill 6868 on June 30 and July 1, 2018 at $\sqrt{s}$ = 13 TeV
\item Zero-bias triggers on 5 bunch pairs (BCIDs 265, 865, 1780, 2192, and 3380)
recorded events rate is 27.7 kHz. 
\item "lsc1", was the "constant separation" scan, two beams were separated by 1.4 $\sigma$ and moved together in steps of 1 $\sigma$ across and back.
\item "lsc2", was the "variable separation" scan method, one beam (starting with beam 1) is moved to -2.5 $\sigma$ and then a three-point scan (a "miniscan") is performed with the other beam.
\item a normal VdM scan pair "norm4" and two short emittance scan pairs "emit4" and "emit5" were also performed.
\end{itemize}

To measure the luminosity using the pixel detector, pixel cluster count is studied for each module, where a module consists of a number of pixels arranged in a grid. The number of pixel clusters recorded in each module per unit time is proportional to the instantaneous luminosity of the collider. To implement PCC method, the pixel detector is operated in an integrated mode, where the detector is read out at a fixed time interval called a "lumi section". During each lumi section, the number of pixel clusters produced in each module is recorded, the rate of pixel clusters and visible cross section of the detector is used to determine the instantaneous luminosity of the collider. PCC data is reprocessed, a procedure that involves the re-analysis of the data using refined or updated algorithms, calibrations, or detector conditions.

\newpage
The purpose of reprocessing can be multifold:

\item Improved Algorithms or Calibration: As experiments progress, there can be improvements in algorithms or calibration techniques used to process the raw data. This can lead to a more accurate or efficient representation of the collected data. Reprocessing allows these improvements to be applied to previously collected data.

\item Changes in Detector Conditions: Detector conditions might change over time due to factors like aging or upgrades. These changes can affect the interpretation of the data. By reprocessing, the data can be reinterpreted to account for these changes.

\item Discovery of Issues or Errors: Sometimes, errors or issues in the original data processing might be discovered. Reprocessing the data allows these errors to be corrected.

In the context of PCC data, reprocessing can lead to a more accurate luminosity measurement, which in turn can affect the interpretation of the physics results derived from the LHC data. The luminosity measurement is crucial because it is directly linked to the determination of cross-sections, a fundamental quantity in the interpretation of the outcomes of high-energy particle physics experiments.

CMS software (CMSSW) version $10{\_}2{\_}2$ is used for reprocessing random trigger PCC dataset and $10{\_}6{\_}30$  \cite{CMSReleaseNotes} is used for reprocessing zero bias PCC datasets. Datasets can be obtained by applying triggers on all bunches (colliding and non colliding) and only on colliding bunches to get random trigger and zero bias data respectively as described in previous section. Random trigger PCC dataset is reprocessed with new module vetolist to obtain afterglow corrections with improved module stability selections. Zero bias dataset is reprocessed after applying afterglow corrections using same module vetolist with more stringent selections on module stability to obtain PCC luminosity.

%To ensure accurate luminosity measurements, it is necessary to correct for various effects, such as detector inefficiencies and pileup (the presence of additional proton-proton collisions in the same bunch crossing). This is achieved through careful calibration of the detector and by using dedicated algorithms to estimate the contribution of pileup to the observed pixel cluster counts.

@@

\begin{figure}[!htp]
\centering
\includegraphics[width=0.8\textwidth]{ashish_thesis/pixel_layer_disk_noveto.png}
\caption{%                                                                                                                                                                
   Luminosity fraction of pixel detector for various layer and disks as a function of lumi section after removing low statistics lumi section.
}
\label{fig:PCC_stab_begin}
\end{figure}

@@

\item Poor statistics lumi sections are removed by applying selection on total PCC Fig. \ref{fig:PCC_cut}. % and \ref{fig:f_it}.

>>>>

For studying the performance of PCC luminometer during 2018 data taking, we have divided CMS luminosity data taking into seven periods: A, B, C, D1, D2, D3, and D4, as depicted in % 2018 CMS luminosity data taking is divided into seven periods namely A, B, C, D1, D2, D3 and D4
Fig. \ref{fig:period_bound}. Run ranges for each period are shown in Table \ref{tab:period_run_ranges}.
During period B, specifically corresponding to Fill 6868, van der Meer (vdM) scans were conducted for calibrating the PCC luminometer.
%vdM data is shown during period B corresponding to Fill 6868.

\begin{figure}[!htp]
\centering
\includegraphics[width=0.9\textwidth]{ashish_thesis/period_boundary.png}
\caption[2018 CMS Luminosity Data Period]{%
   2018 luminosity data taking periods showing data taking period boundaries and vdM calibration data (1 July in period B)  \cite{CERNLumiPublicResults}.
}
\label{fig:period_bound}
\end{figure}

\begin{table}
  \begin{center}
    \caption[Run ranges for 2022 periods]{2018 luminosity data periods with run ranges.}
    \begin{tabular}{ccccc}
    \textbf{Period}   & \textbf{Run range} \\ \hline
     2018A      &        315252-316995         \\
     2018B      &        317080-319311         \\
     2018C      &        319337-320065         \\
     2018D1     &        320500-321665        \\
     2018D2     &        321710-322964         \\
     2018D3     &        323363-324420         \\
     2018D4     &        324564-325175        \\
      \end{tabular}
    %\caption[Run ranges]{2018 luminosity data periods with run ranges.}
    \label{tab:period_run_ranges}
  \end{center}
\end{table}

In the study under consideration, the datasets have been drawn from the Alignment and Calibration (AlCa) project, an integral component of the CMS experiment. The dataset selection includes two subsets: Random trigger and Zero-Bias. The Random trigger dataset comprises of data where trigger is applied on both colliding and non-colliding bunches. On the other hand, the Zero-Bias dataset encapsulates events that have passed the trigger system with no bias which means the colliding bunches are randomly triggered without the use of any CMS detector signal. The list of colliding bunches is detected by an independent system Beam Pick-up Timing system (BPTX) for Experiments. CMS software (CMSSW) version $10{\_}2{\_}2$ is used for reprocessing random trigger PCC dataset and $10{\_}6{\_}30$  \cite{CMSReleaseNotes} is used for reprocessing zero bias PCC datasets.

@@

The calibration of luminosity measurement by the CMS experiment luminometers in the 2018 proton-proton data taking at $\sqrt{s}$ = 13 TeV \cite{CMS-PAS-LUM-18-002} was performed during LHC fill 6868 on June 30 and July 1, 2018 at $\sqrt{s}$ = 13 TeV. Zero-bias triggers on 5 bunch pairs (BCIDs 265, 865, 1780, 2192, and 3380) recorded events rate is 27.7 kHz. Trigger is applied on five colliding bunches due to limitation of the recording bandwidth.

%\item "lsc1", was the "constant separation" scan, two beams were separated by 1.4 $\sigma$ and moved together in steps of 1 $\sigma$ across and back.
%\item "lsc2", was the "variable separation" scan method, one beam (starting with beam 1) is moved to -2.5 $\sigma$ and then a three-point scan (a "miniscan") is performed with the other beam.
%a normal VdM scan pair "norm4" and two short emittance scan pairs "emit4" and "emit5" were also performed.

%The 2018 CMS VdM scan program was conducted in two segments, interrupted by an alarm. The first segment involved six x-y scan pairs. Two were short ``emittance'' scans, named ``emit1'' and ``emit2,'' where the beams were separated by $ \(4\sigma_b\) $ over 9 steps with a 10 second integration time at each step. A standard VdM scan, named ``norm1,'' separated the beams by $\(6\sigma_b \approx 600\)~\mu m$ in 25 steps, with a 30-second duration per step. The ``offset1'' scan followed the same procedure as the standard scans but with a $\(\pm 1.5\sigma_b\)$ separation in the non-scanning direction. Lastly, two sets of ``beam imaging'' scans were conducted, where one beam remained fixed while the other was moved in 19 steps between $\(+4.5\sigma_b\)$ and $\(-4.5\sigma_b\)$, each lasting 46 seconds per step.

The 2018 CMS VdM scan program (shown in Fig. \ref{fig:vdm_prog_2018}) was conducted in two segments, interrupted by an alarm. The first segment involved six x-y scan pairs. Two were short ``emittance'' scans, named ``emit1'' and ``emit2,'' where the beams were separated by \(4\sigma_b\) over 9 steps with a 10-second integration time at each step. A standard vdM scan (described in section 3.1), named ``norm1,'' separated the beams by \(6\sigma_b \approx 600\)~$\mu$m in 25 steps, with a 30-second duration per step. The ``offset1'' scan followed the same procedure as the standard scans but with a \(\pm 1.5\sigma_b\) separation in the non-scanning direction. Lastly, two sets of ``beam imaging'' scans were conducted, where one beam remained fixed while the other was moved in 19 steps between \(+4.5\sigma_b\) and \(-4.5\sigma_b\), each lasting 46 seconds per step. The second part of the scan program was conducted in the same fill, approximately 7.5 hours later, and consisted of twelve scan pairs. This included a short emittance scan, ``emit3''; beam imaging scan pairs, ``imag2'' and ``imag3''; an offset scan pair, ``offset2''; and two standard VdM scan pairs, ``norm2'' and ``norm3''.

Beam Imaging Scans: Beam imaging scans are designed to measure the transverse profile of the beam. They involve moving one beam across the other in steps and measuring the interaction rate (or event rate) at each step. By doing this, the shape or "profile" of the beam in the transverse plane can be obtained. These scans allow for a detailed analysis of the beam shape, which is essential for understanding the beam properties. In a perfect scenario, the density distributions of protons in a bunch would factorize into separate X (horizontal) and Y (vertical) components. However, real beams can have correlations between X and Y (non-factorization). Beam imaging scans help map the 2D profile of the beam and are thus essential for diagnosing and understanding any such correlations. Knowing the 2D profile of the beams is crucial when calculating luminosity, as assumptions made about the factorization of the beam profiles feed into these calculations. By using beam imaging scans, experimenters can assess the validity of the factorization assumption or correct for any observed non-factorization.

Offset Scans: Offset scans are performed to check for systematic errors in the luminosity measurement, which may arise from various sources. They are similar to standard VdM scans but involve an intentional offset in the non-scanning direction. By comparing the results of offset scans with the results of regular scans, one can identify and correct for any systematic discrepancies that are present. Offset scans can help to assess the sensitivity of the luminosity measurement to non-factorization in the transverse plane. For example, if there are correlations between the X and Y distributions of protons (i.e., non-factorization), this could in principle affect the shape of the beam overlap region in the collision point. Offset scans, by changing the relative position of the beams, can help to assess whether such non-factorization has a significant effect on the measured luminosity.

%The second part of the scan program was conducted in the same fill, approximately 7.5 hours later, and consisted of twelve scan pairs. This included a short emittance scan, ``emit3''; beam imaging scan pairs, ``imag2'' and ``imag3''; an offset scan pair, ``offset2''; and two standard VdM scan pairs, ``norm2'' and ``norm3''.

vdM scans are typically conducted at least once a year to ensure the luminometer is correctly calibrated. Results of these scans that is the visible cross section is used to normalize the data collected by the experiments during physics run.

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{ashish_thesis/vdm_program_2018.png}
    \caption[2018 vdM program]{Beam position along x and y is plotted as a function of time showing different types of scans. #3 and #10 are vdM scans, #4 and #9 are offset scans, #5, #7 and #8 are beam imaging scans, #6 is emittance scan.
}
    \label{fig:vdm_prog_2018}
\end{figure}

%In order to measure luminosity using a pixel detector, the count of pixel clusters is studied for each module, with each module being composed of numerous pixels arranged in a grid. The number of pixel clusters recorded per unit time in each module is proportional to the instantaneous luminosity of the collider. The Pixel Cluster Counting (PCC) method is implemented by operating the pixel detector in an integrated mode. The Pixel detector is read out at a fixed interval termed a "lumi section", each lasting for 23.36 seconds.

The pixel detector records data for each bunch-crossing, though the readout rate is constrained by the trigger. During subsequent data processing,  these events are consolidated into lumi sections. During each lumi section, the number of pixel clusters produced in each module is recorded and the module weights for all pixel modules are studied to track their performance over time. This helps in identifying good and bad modules, as shown in Fig. \ref{fig:goodbadmodules}. Only pixel modules that demonstrate consistent and reliable performance throughout the entire 2018 year are considered good and included in the analysis. The pixel cluster rate, along with the visible cross section of the detector, is used to determine the instantaneous luminosity of the collider. PCC data is subsequently reprocessed. This involves the re-analysis of the data using updated or refined algorithms, calibrations, and detector conditions.

%To measure the luminosity using the pixel detector, pixel cluster count is studied for each module, where a module consists of a number of pixels arranged in a grid. The number of pixel clusters recorded in each module per unit time is proportional to the instantaneous luminosity of the collider. To implement PCC method, the pixel detector is operated in an integrated mode, where the detector is read out at a fixed time interval called a "lumi section". During each lumi section, the number of pixel clusters produced in each module is recorded, the rate of pixel clusters and visible cross section of the detector is used to determine the instantaneous luminosity of the collider. PCC data is reprocessed, a procedure that involves the re-analysis of the data using refined or updated algorithms, calibrations, or detector conditions.

The purpose of reprocessing can be multifold:

@@

\item Improved Algorithms or Calibration: As experiments progress, there can be improvements in algorithms or calibration techniques used to process the raw data. This can lead to a more accurate or efficient representation of the collected data.

\item Changes in Detector Conditions: Detector conditions might change over time due to factors like aging or upgrades. These changes can affect the interpretation of the data.

\item Discovery of Issues or Errors: Sometimes, errors or issues in the original data processing might be discovered.

@@

In the context of PCC data, reprocessing can lead to a more accurate luminosity measurement, which in turn can affect the interpretation of the physics results derived from the LHC data. Random trigger PCC dataset is reprocessed with new module veto list (described in section 4.2) to obtain afterglow corrections with improved module stability selections. Zero-bias dataset is reprocessed after applying afterglow corrections using same module veto list with more stringent selections on module stability to obtain PCC luminosity.

%CMS software (CMSSW) version $10{\_}2{\_}2$ is used for reprocessing random trigger PCC dataset and $10{\_}6{\_}30$  \cite{CMSReleaseNotes} is used for reprocessing zero bias PCC datasets. %Datasets can be obtained by applying triggers on all bunches (colliding and non colliding) and only on colliding bunches to get random trigger and zero bias data respectively as described in previous section.
%Random trigger PCC dataset is reprocessed with new module veto list (described in section 4.2) to obtain afterglow corrections with improved module stability selections. Zero-bias dataset is reprocessed after applying afterglow corrections using same module veto list with more stringent selections on module stability to obtain PCC luminosity. 
%To ensure accurate luminosity measurements, it is necessary to correct for various effects, such as detector inefficiencies and pileup (the presence of additional proton-proton collisions in the same bunch crossing). This is achieved through careful calibration of the detector and by using dedicated algorithms to estimate the contribution of pileup to the observed pixel cluster counts.

@@

%\begin{figure}[!htp]
%\centering
%\includegraphics[width=0.8\textwidth]{ashish_thesis/pixel_layer_disk_noveto.png}
%\caption{%                                                                                                                                                                
 %  Luminosity fraction of pixel detector for various layer and disks as a function of lumi section after removing low statistics lumi section.
%}
%\label{fig:PCC_stab_begin}
%\end{figure}

@@

\item Poor statistics lumi sections are removed by applying selection on total PCC. %Fig. \ref{fig:PCC_cut}. % and \ref{fig:f_it}.
